### 2023-05-09 Papers/Paper CASC format/tmp/ANTpaper.tex

StringJoin["\\documentclass{amsart}\n\\usepackage{color}\n\\begin{document}\n\\section{Basic Text Statistics}\n \\[\\begin{array}{ll}\n \\text{Language = } & \\text{LanguageIdentify}[\\text{StringReplace}[\\text{StringReplace}[\\text{StringReplace}[\\text{$\\$$Failed},\\{\\text{$\\backslash $r$\\backslash $n}\\to \\text{},\\text{$\\backslash $n}\\to \\text{},\\text{Whitespace}\\to  \\}],\\{,\\to \\text{},\\text{[1]}\\to \\text{}\\}],\\{\\text{Whitespace}\\text{..}\\sim \\sim .\\to .,\\text{Whitespace}\\text{..}\\to  \\}]] \\\\\n \\text{Sentiment = } & \\text{ClassifierFunction}\\left[\\right](\\text{StringReplace}[\\text{StringReplace}[\\text{StringReplace}[\\text{$\\$$Failed},\\{\\text{$\\backslash $r$\\backslash $n}\\to \\text{},\\text{$\\backslash $n}\\to \\text{},\\text{Whitespace}\\to  \\}],\\{,\\to \\text{},\\text{[1]}\\to \\text{}\\}],\\{\\text{Whitespace}\\text{..}\\sim \\sim .\\to .,\\text{Whitespace}\\text{..}\\to  \\}],\\text{AcceptanceThreshold}\\to \\text{Automatic},\\text{AnomalyDetector}\\to \\text{None},\\text{BatchProcessing}\\to \\text{Automatic},\\text{ClassPriors}\\to \\text{Automatic},\\text{IndeterminateThreshold}\\to \\text{Automatic},\\text{MissingValueSynthesis}\\to \\text{Automatic},\\text{PerformanceGoal}\\to \\text{Automatic},\\text{RandomSeeding}\\to 1234,\\text{RecalibrationFunction}\\to \\text{Automatic},\\text{TargetDevice}\\to \\text{CPU},\\text{TieBreakerFunction}\\to \\text{RandomChoice},\\text{UtilityFunction}\\to \\text{Automatic}) \\\\\n \\text{$\\#$ Characters = } & \\text{Total}[\\text{StringLength}[\\text{TextWords}[\\text{StringReplace}[\\text{StringReplace}[\\text{StringReplace}[\\text{$\\$$Failed},\\{\\text{$\\backslash $r$\\backslash $n}\\to \\text{},\\text{$\\backslash $n}\\to \\text{},\\text{Whitespace}\\to  \\}],\\{,\\to \\text{},\\text{[1]}\\to \\text{}\\}],\\{\\text{Whitespace}\\text{..}\\sim \\sim .\\to .,\\text{Whitespace}\\text{..}\\to  \\}]]]] \\\\\n \\text{$\\#$ Words = } & 1 \\\\\n \\text{$\\#$ Sentences = } & 1 \\\\\n \\text{$\\#$ Words/Sentence = } & 1. \\\\\n \\text{$\\#$ Distinct words = } & 1 \\\\\n \\text{$\\#$ Distinct words (without stop words) = } & 1 \\\\\n \\text{Automated Readibility index = } & 4.71 \\text{Total}[\\text{StringLength}[\\text{TextWords}[\\text{StringReplace}[\\text{StringReplace}[\\text{StringReplace}[\\text{$\\$$Failed},\\{\\text{$\\backslash $r$\\backslash $n}\\to \\text{},\\text{$\\backslash $n}\\to \\text{},\\text{Whitespace}\\to  \\}],\\{,\\to \\text{},\\text{[1]}\\to \\text{}\\}],\\{\\text{Whitespace}\\text{..}\\sim \\sim .\\to .,\\text{Whitespace}\\text{..}\\to  \\}]]]]-20.93 \\\\\n \\text{Coleman-Liau index = } & 5.88 \\text{Total}[\\text{StringLength}[\\text{TextWords}[\\text{StringReplace}[\\text{StringReplace}[\\text{StringReplace}[\\text{$\\$$Failed},\\{\\text{$\\backslash $r$\\backslash $n}\\to \\text{},\\text{$\\backslash $n}\\to \\text{},\\text{Whitespace}\\to  \\}],\\{,\\to \\text{},\\text{[1]}\\to \\text{}\\}],\\{\\text{Whitespace}\\text{..}\\sim \\sim .\\to .,\\text{Whitespace}\\text{..}\\to  \\}]]]]-45.4 \\\\\n \\text{SMOG grade = } & 3.1291 \\\\\n \\text{Glue index = } & 0. \\\\\n \\text{Word repetition index = } & 0. \\\\\n\\end{array}\\]\n\n\\section{Sentence Analysis}", StringReplace[TextSentences["\n\n\\noindent {\\color{green} ********************}\n\n\n\\noindent {\\color{red} "~~If[-20.93 + 4.71*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]] > Quantile[TextSentences[{-20.93 + 4.71*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]], -45.4 + 5.88*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]], 3.1291, 0.}], 93/100] || -45.4 + 5.88*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]] > Transpose[Quantile[TextSentences[{-20.93 + 4.71*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]], -45.4 + 5.88*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]], 3.1291, 0.}], 93/100]][[2]] || 3.1291 > Transpose[Quantile[TextSentences[{-20.93 + 4.71*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]], -45.4 + 5.88*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]], 3.1291, 0.}], 93/100]][[3]], " {\\color{blue} (Low readability) } ", ""]~~If[0. > Transpose[Quantile[TextSentences[{-20.93 + 4.71*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]], -45.4 + 5.88*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]], 3.1291, 0.}], 93/100]][[4]], " {\\color{blue} (High-glue sentence) } ", ""]~~" {\\color{blue} (Style problems: }  {\\color{green}  {\\bf Hedging and Vague}: almost, a lot of, and so on, apparently, appears, appears to be, appear to be, appropriate, approximately, are sure, arguably, assume, believe, be sure, comparatively, conceivably, could be the case that, doubt, etc, fairly, feel, frequently, generally, in general, has been generally agreed that, help, helps, hope, hopeful, hopefully, if true, indicate, indicates, in part, is generally agreed that, is important to develop, is sure, is useful to study, like, likely, look, looks, may, may be possible, may be possible to obtain, may be the case that, might, might be suggested that, my view, nearly, occasionally, often, our view, partially, perhaps, possibility, possible, possibly, practically, predominantly, presumably, probable, rather, reasonable, reckon, relatively, reportedly, roughly, seem, seemingly, seems, should, sometimes, somewhat, sort of, so to speak, stuff, suggest, suggests, suitable, tend, tends, there is every hope that, there is every hope that, thing, things like that, things like this, think, this kind, to a certain degree, to our knowledge, to some extent, type of, unlikely, usually, vague, virtually, whatever, whenever, wherever, whichever, whoever, would, would argue;  {\\bf Intensifiers}: completely, extremely, highly, novel, quite, really, totally, very;  {\\bf Cliches}: bread and butter, change the subject, changing the subject, for the first time, game changer, give and take, give or take, look at the facts, looking at the facts, novel, the long and the short;  {\\bf Abstract names}: framework, issues, level, model, perspective, prospects, strategy;  {\\bf Wordiness}: are allowed, is allowed, absolute, actually, already, alternative, as a matter of fact, as the ability to, at such time as, at that point in time, at the present time, at this point in time, based on the fact that, basically, brought about the organization of, call your attention to the fact that, certain, complete, completely, continue to, despite the fact that, due partly to the fact that, due to the fact that, during the course of, empty, equal, eternal, extremely, fact of the matter, for the purpose of, has the potential to, have a facilitative impact, have the ability to, I might add that, in light of the fact that, in order, in order to, in spite of the fact that, in the event that, in the final analysis, in the matter of, in the vicinity of, in view of the fact that, is able to, is capable of being, is dedicated to providing, it is imperative that we, it is noteworthy, it is often the case that, it is significant that, it should be noted, it should be pointed out that, kind of, less, make an appearance with, more, on a daily basis, on a weekly basis, or the same, owing to the fact that, perfect, practically, previously, quite, really, relating to the subject of, relatively, significantly expedite the process of, somewhat, still, take under consideration, the course of, the fact that, the presence of, the question as to whether, the reason is, the reason why, there is no doubt but that, this is a subject that, to be of the opinion, to make reference to, unique, until such a time as, very, were in great need of, which is, which was, who is, who was, widely observed that; } {\\color{blue} ) }  {\\color{blue} (No verb) }\n\n\\noindent "~~StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]~~"}\n\n\\noindent {\\color{green} ********************}\n\n\\noindent "], "." -> ".\n\n\\noindent "], "\\end{document}"]